Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-02-23 local-capability-machine.tex
Added another sanity lemma
fixed ref issues (by reinserting small lemma in proof that was removed by mistake)

To show
\[
  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_\stk+1,\addrend_\stk})} \in \observations(W)
\]
We use Lemma~\ref{lem:scall-works} (we do not use the local frame in the lemma) which requires us to show
\begin{enumproof}
  \item $\memSat[n-i]{\ms'}{W}$ \\
    Partition $\ms'$ as follows:
    \begin{enumproof}
      \item $\ms_\malloc$: governed by $\iota_{\malloc,0}$, use malloc specification.
      \item $\ms_\flag \uplus \ms_{f2}$: governed by region 1, only this memory segment is accepted.
      \item $\ms_\link$: governed by region 2, only this memory segment is accepted. We also need to show that the contents is safe, i.e. shoe
        \begin{enumproof}
          \item $\npair[n-i]{c_\malloc} \in \stdvr(W)$: Follows from Lemma~\ref{lem:malloc-in-vr}.
          \item $\npair[n-i]{c_\adv} \in \stdvr(W)$: Follows from Theorem~\ref{thm:ftlr} and Lemma~\ref{lem:nwlp-subset-pwl}.
        \end{enumproof}
      \item $\ms_\adv$:
        Follows from Lemma~\ref{lem:stau-subset-pwl}.
    \end{enumproof}
  \item Hyp-Callee \\
    Assume
    \begin{itemize}
    \item $\dom(\ms_\stk |_{\start_\stk+1,\addrend_\stk}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}')$
    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1]\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
    \item $\memSat[n-i-1]{\ms''}{W'}$
    \item $\reg'' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
    \item $\reg''= \reg_0[\pcreg\mapsto\updatePcPerm{\reg'(r)},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk', r \mapsto \reg'(r)]$ 
    \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$
    \item $\npair[n-i-1]{c_\stk'} \in \stdvr(W')$
    \end{itemize}
    for some $\ms_{\mathit{act}}$, $\ms_{\mathit{unused}}$, $\ms''$, $\reg''$, $c_{\mathit{ret}}$.

@@

\begin{lemma}
  \label{lem:pwl-stack}
  For all $W$, $\start$, $\addrend$, $n$, $\ms$ if
  \begin{itemize}
  \item $\memSat{\ms}{W}$
  \item $\npair{((\rwlx,\gl),\start,\addrend,\addr)} \in \stdvr(W)$
  \item $\start \leq \addrend$
  \end{itemize}
  then
  \[
    \exists r, \start',\addrend' \ldotp [\start,\addrend] \subseteq [\start',\addrend'] \land W(\var{r}) \nequal \iota^\pwl_{\start',\addrend'}
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:pwl-stack}]
  Assume
  \begin{enumproof}
    \item $\npair{((\rwlx, \gl),b,e,a)} \in \stdvr(W)$ \label{ass:pwl-stack:val-rel}
    \item $\memSat{\ms}{W}$ \label{ass:pwl-stack:mem-sat}
  \end{enumproof}
  From Assumption~\ref{ass:pwl-stack:val-rel}, we get $r_1$, $r_2$, $b_1$, $b_2$, $e_1$ and $e_2$ such that
  \begin{enumproof}[resume]
    \item $r_1 \in \localityReg(\gl,W)$
    \item $r_2 \in \localityReg(\gl,W)$
    \item $[b,e] \subseteq [b_1,e_1]$\label{ass:pwl-stack:b1}
    \item $[b,e] \subseteq [b_2,e_2]$ \label{ass:pwl-stack:b2}
    \item $W(r_1) \nsubsim \iota^\pwl_{b_1,e_1}$ \label{ass:pwl-stack:subsim}
    \item $W(r_2) \nsupsim \iota^\pwl_{b_2,e_2}$ \label{ass:pwl-stack:supsim}
    \item $W(r_2)$ is address-stratified. \label{ass:pwl-stack:add-stra}
  \end{enumproof}
From Assumption~\ref{ass:pwl-stack:mem-sat}, we get partitionen $P$ s.t.
\[
  \memSat[n,p]{\ms}{W}
\]
Say $P(r_1) = \ms_1$ and $P(r_2) = \ms_2$. First from $\npair{\ms_1} \in W(r_1).H\;W(r_1).s\;\xi^{(-1)}(W)$ using \label{ass:pwl-stack:subsim}, we get $\npair{\ms_1} \in H^\pwl_{b_1,e_1}\;1\;\xi^{(-1)}(W)$ which means $\dom(\ms_1) = [b_1,e_1]$.

@@

>>>>

To show
\[
  \npair[n-i]{(\reg',\ms' \uplus [\start_\stk \mapsto 1] \uplus \ms_\stk |_{\start_\stk+1,\addrend_\stk})} \in \observations(W)
\]
We use Lemma~\ref{lem:scall-works} (we do not use the local frame in the lemma) which requires us to show
\begin{enumproof}
  \item $\memSat[n-i]{\ms'}{W}$ \\
    Partition $\ms'$ as follows:
    \begin{enumproof}
      \item $\ms_\malloc$: governed by $\iota_{\malloc,0}$, use malloc specification.
      \item $\ms_\flag \uplus \ms_{f2}$: governed by region 1, only this memory segment is accepted.
      \item $\ms_\link$: governed by region 2, only this memory segment is accepted. We also need to show that the contents is safe, i.e. shoe
        \begin{enumproof}
          \item $\npair[n-i]{c_\malloc} \in \stdvr(W)$: Follows from Lemma~\ref{lem:malloc-in-vr}.
          \item $\npair[n-i]{c_\adv} \in \stdvr(W)$: \\
            We will show
            \begin{equation}
              \label{eq:f3:adv-in-stder}
              \forall W' \futurestr W \ldotp \npair{c_\adv} \in \stdvr(W')
            \end{equation}
            which will give us what we need using downwards closure as well as a result for later use.

            Let $W' \futurestr W$ be given and show
            \[
              \npair{(\start_\adv,\addrend_\adv,\start_\adv+\olf)} \in \entryCond{}(\glob)(W')
            \]
            to this end let $W'' \futurestr W'$ and $n' < n$ be given and show
            \[
              \npair[n']{\updatePcPerm{c_\adv}} \in \stder(W'')
            \]
            This follows from the FTLR (Theorem~\ref{thm:ftlr}) if we can show
            \[
              \npair[n']{\start_\adv,\addrend_\adv} \in \readCond{}(\glob)(W'')
            \]
            $\iota^{\nwl,p}_{\start_\adv,\addrend_\adv}$ governs the adversary, so the result follows from Lemma~\ref{lem:nwlp-subset-pwl}.
          \end{enumproof}
      \item $\ms_\adv$:
        Follows from Lemma~\ref{lem:stau-subset-pwl}.
    \end{enumproof}
  \item Hyp-Callee \\
    Assume
    \begin{itemize}
    \item $\dom(\ms_\stk |_{\start_\stk+1,\addrend_\stk}) = \dom(\ms_{\mathit{act}} \uplus \ms_{\mathit{unused}}')$
    \item $W' = \revokeTemp{W}[\iota^{\sta}(\temp,[\start_\stk \mapsto 1]\uplus\ms_{\mathit{act}}),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$
    \item $\memSat[n-i-1]{\ms''}{W'}$
    \item $\reg'' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
    \item $\reg''= \reg_0[\pcreg\mapsto\updatePcPerm{\reg'(r)},r_0 \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk', r \mapsto \reg'(r)]$ 
    \item $\npair[n-i-1]{c_{\mathit{ret}}} \in \stdvr(W')$
    \item $\npair[n-i-1]{c_\stk'} \in \stdvr(W')$
    \end{itemize}
    for some $\ms_{\mathit{act}}$, $\ms_{\mathit{unused}}$, $\ms''$, $\reg''$, $c_{\mathit{ret}}$.

@@

\begin{lemma}
  \label{lem:pwl-stack}
  For all $W$, $\start$, $\addrend$, $n$, $\ms$ if
  \begin{itemize}
  \item $\memSat{\ms}{W}$
  \item $\npair{\stdcap} \in \stdvr(W)$
  \item $\start \leq \addrend$
  \item $\perm \in \{ \rwlx, \rwx \}$
  \end{itemize}
  then
  \[
    \exists r, \start',\addrend' \ldotp [\start,\addrend] \subseteq [\start',\addrend'] \land W(\var{r}) \nequal \iota^\pwl_{\start',\addrend'}
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:pwl-stack}]
  Assume
  \begin{enumproof}
    \item $\npair{((\rwlx, \gl),b,e,a)} \in \stdvr(W)$ \label{ass:pwl-stack:val-rel}
    \item $\memSat{\ms}{W}$ \label{ass:pwl-stack:mem-sat}
  \end{enumproof}
  From Assumption~\ref{ass:pwl-stack:val-rel}, we get $r_1$, $r_2$, $b_1$, $b_2$, $e_1$ and $e_2$ such that
  \begin{enumproof}[resume]
    \item $r_1 \in \localityReg(\gl,W)$
    \item $r_2 \in \localityReg(\gl,W)$
    \item $[b,e] \subseteq [b_1,e_1]$\label{ass:pwl-stack:b1}
    \item $[b,e] \subseteq [b_2,e_2]$ \label{ass:pwl-stack:b2}
    \item $W(r_1) \nsubsim \iota^\pwl_{b_1,e_1}$ \label{ass:pwl-stack:subsim}
    \item $W(r_2) \nsupsim \iota^\pwl_{b_2,e_2}$ \label{ass:pwl-stack:supsim}
    \item $W(r_2)$ is address-stratified. \label{ass:pwl-stack:add-stra}
  \end{enumproof}
From Assumption~\ref{ass:pwl-stack:mem-sat}, we get partitionen $P$ s.t.
\[
  \memSat[n,p]{\ms}{W}
\]
Say $P(r_1) = \ms_1$ and $P(r_2) = \ms_2$. First from $\npair{\ms_1} \in W(r_1).H\;W(r_1).s\;\xi^{(-1)}(W)$ using \label{ass:pwl-stack:subsim}, we get $\npair{\ms_1} \in H^\pwl_{b_1,e_1}\;1\;\xi^{(-1)}(W)$ which means $\dom(\ms_1) = [b_1,e_1]$.

@@

\begin{lemma}
  \label{lem:wl-local}
  If
  \begin{itemize}
  \item $\memSat{\ms}{W}$
  \item $\npair{\stdcap} \in stdvr(W)$
  \item $\start \leq \addrend$
  \item $\perm \in \{\rwlx,\rwl\}$
  \end{itemize}
  then 
  \[
    \gl = \local
  \]
\end{lemma}
\begin{proof}[Proof of Lemma~\ref{lem:wl-local}]
It follows as a consequence of Lemma~\ref{lem:pwl-stack}. The $n$-equality forces the region to be $\temp$, so for the region name to be in $\localityReg(\gl,W)$, the locality must be $\local$.
\end{proof}
